---
title: "Inferences About Population Variances"
author: "Ramsey El Lethy"
date: "2023-10-07"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Bus Times}
#read our files
bus_times_df <- read.csv("bustimes_r.csv")

#use the function nrow to gather our sample size

samp_size <- nrow(bus_times_df) #24 observations 

#creating a vector (array) of times 
times <- bus_times_df$Times

#using the R funciton var to calculate the
#sample variance of the variable 

var_time <- var(times) #4 .9 (sample varaince)

#Calculate our Chi Squared Stat 

test_chi <- (samp_size - 1) * var_time / 4

#pchisq gives us the p- value 

p_value <- pchisq(test_chi,samp_size-1,lower.tail = FALSE)

list(test_chi, p_value)
```
This test gives us a Chi Squared Value of 28.18 
and a p-value of 0.209

We are comparing the p-value to an alpha value of 0.05, since
our p-value is greater, we fail to reject the null hypothesis, 
We do not have significant data to suggest the variance is greater than 4 

```{r Bus Types}
#Two Tailed Test 

#gather our data
school_bus_df <- read.csv("schoolbus_r.csv")

#Create subsets separating the vars
milbank_df <- subset(school_bus_df, select = Milbank)
gulfpark_df <- subset(school_bus_df, select = Gulf.Park)

#omit the rows with missing values
milbank_df <- na.omit(milbank_df)
gulfpark_df <- na.omit(gulfpark_df)

#gather the sample size
samp_size_milbank <- nrow(milbank_df)
samp_size_gulfpark <- nrow(gulfpark_df)

#create a vector of times for milbank and gulf park 
milbank_times <- milbank_df$Milbank
gulfpark_times <- gulfpark_df$Gulf.Park

#var to calculate sample variance
var_time_milbank <- var(milbank_times)
var_time_gulfpark <- var(gulfpark_times)

#We use max and min functions to calculate F stat
test_f <- max(var_time_milbank, var_time_gulfpark) /
  min(var_time_milbank, var_time_gulfpark)

#We use an If-else to find the numerator and denom
#degrees of freedom 

if (var_time_milbank > var_time_gulfpark) {
  degreesnum = samp_size_milbank -1;
  degreesdenom = samp_size_gulfpark - 1;
} else {
  degreesnum = samp_size_gulfpark -1; 
  degreesdenom = samp_size_milbank - 1; 
}

#pf to perform hypothesis test  (two tailed, thats why we * 2)

p_value_bus <- 2 * pf(test_f, degreesnum, degreesdenom, lower.tail = FALSE)

list(test_f,p_value)
```
We get a test statistic of 2.401, with degrees freedom of 15 in the
numerator and 25 in the denominator, we get a p-value of 0.081
comparing this to an alpha value of 0.05, 0.081 > 0.05, so we 
fail to reject the null hypothesis, there is not enough evidence
to suggeset the variances are different. 
```{r Vehicle Ownership}
#vehicle ownership by Fortune Magazine Subscribers 
# Ho: o = 0.94 Ha: o != 0.94 (Two tailed test)

#vector of data
vehicle <- c(2,1,2,0,3,2,2,1,2,1,0)

#turn that into a data frame
vehicle_df <- data.frame(vehicle)

#gather the sample size

sample_vehicle <- nrow(vehicle)

#calculate sample variance

var_vehicle <- var(vehicle)

#Calculate our Chi Squared Statistic 

test_chi_vehicle <- (sample_vehicle - 1) * var_vehicle / 0.94

#pchisq p value for 1 tail, * 2 for 2 tail 

p_value_vehicle <- 2 * pchisq(test_chi,samp_size-1,lower.tail = FALSE)

list(var_vehicle, test_chi_vehicle, p_value_vehicle)
```
A. Sample Variance = 0.8727

B. Our Chi squared Statistic was calculated to be 9.28 and 
our p-value was calculated to be 1.98. Comparing it to an alpha 
value of 0.05. 1.98 > 0.05, so we fail to reject the null hypothesis
there is not enough evidence to conclude that the Population variance is
not equal to 0.94. 

```{r Bag Filling Machines }
#Bag-Filling Machines at Jelly belly
#Test of equality between to populatins 
#H0: o1 = o2
#HA: o1 != o2
#ASSUME THE POPULATION VARIANCES ARE EQUAL IF 
#H0 IS REJECTED WE DRAW THE CONCLUSION THAT THE 
#POPULATION VARIANCES ARE NOT EQUAL 

#gather our data 
bags_df <- read.csv("bags_r.csv")

#create subsets separating our vars
machine1_df <- subset(bags_df, select = Machine1)
machine2_df <- subset(bags_df, select = Machine2)

#omit rows with missing values 
machine1_df <- na.omit(machine1_df)
machine2_df <- na.omit(machine2_df)

#gather sample size
machine1_size <- nrow(machine1_df)
machine2_size <- nrow(machine2_df)

#create a vector of values for both machines
machine1_vals <- machine1_df$Machine1
machine2_vals <- machine2_df$Machine2

#calculate sample variance respectively
var_machine1 <- var(machine1_vals)
var_machine2 <- var(machine2_vals)

#Calculating our f statistic 
test_f_jellybelly <- max(var_machine1, var_machine2) / min(var_machine1, var_machine2)

#An If else statement to choose our numerator and denom dof
if (var_machine1 > var_machine2) {
  dofnum = machine1_size - 1;
  dofdenom = machine2_size -1; 
} else {
  dofnum = machine2_size -1; 
  dofdenom = machine1_size - 1; 
}


#pf to perform hyp test (two tailed)
p_value_jellybelly <- 2 * pf(test_f_jellybelly, dofnum, dofdenom, lower.tail = FALSE)

list(test_f_jellybelly,p_value_jellybelly)
```
This test gives us a F statistic of 8.28 and a p-value of 
7.22 * 10^-6. Comparing this to an alpha value of 0.05, our
p-value is less than our alpha value. So we can reject the null hypothesis
there is enough evidence to conclude that the variation
between the two machines is significant. 
```{r Smartphone Battery Life }
#Smartphone Battery Life 
#Ratio of variances 
#H0 : 01 <= 02
#HA : 01 > 02 01 is talk time 02 is internet applications 

batterylife_df <- read.csv("batterylife_r.csv")

#Create subsets separating the variables 
talk_df <- subset(batterylife_df, select = Talking)
net_df <- subset(batterylife_df, select = Internet)

#Agian, we omit the rows with missing values
talk_df <- na.omit(talk_df)
net_df <- na.omit(net_df)

#gather our sample sizes
talk_size <- nrow(talk_df)
net_size <- nrow(net_df)

#create a vector of battery life between charges
talk_battery <- talk_df$Talking
net_battery <- net_df$Internet

#var to calculate sample variance 
var_talk <- var(talk_battery)
var_net <- var(net_battery)

#Max and Min functions to calculate our F statistic
test_f_battery <- max(var_talk, var_net) / min(var_talk, var_net)

#If else statement to choose our num and denom dof 
if (var_talk > var_net) {
  dofnumbattery = talk_size - 1; 
  dofdenombattery = net_size - 1; 
} else {
  dofnumbattery = net_size - 1; 
  dofnumbatter = talk_size - 1; 
}

#pf to perform hypothesis test (one tailed upper)
p_value_battery <- pf(test_f_battery, dofnumbattery, dofdenombattery, 
                      lower.tail = FALSE)
#calculate the standard deviation for each function
talk_sd <- sd(talk_battery)
net_sd <- sd(net_battery)

list(test_f_battery, p_value_battery, talk_sd, net_sd)
```
A. 
HO: O1 <= O2
HA: O1 > O2

B. The sample standard Deviations were 
Talk time : 7.363 
Internet Applicaiton : 4.76

This test gives us a F statistic of 2.38 and a p-value of 
0.09. Comparing this to an alpha value of 0.05, our
p-value is greater than our alpha value. So we fail reject the null hypothesis
there is not enough evidence to conclude that the variation
in talk time is greater than the variation in internet applications