---
title: "Experimental Design and Anlysis of Variance"
author: "Ramsey EL Lethy"
date: "2023-10-21"
output:
  pdf_document: default
  html_document: default
---


```{r Auditing Errors}

audit_df <- read.csv("audjudg_r.csv")


audit_anova <- aov(Errors ~ Method, data = audit_df)

anova(audit_anova)

```

For this analysis of variance, we have a p value that is much less than an
alpha value of 0.05, so we reject the null hypothesis. There data is 
significant enough to conclude that the basis of judgement affects the 
quality of judgement

```{r Paint Drying Robots}

paint_df <- read.csv("paint_r.csv")

paint_anova <- aov(DryTime ~ Paint, data = paint_df)

anova(paint_anova)

```

For this analysis of variance, we have a p value (0.09) that is greater than an
alpha value of 0.05, so we fail to reject the null hypothesis. The data is not
significant enough to suggest the mean drying time is different for each brand
of robot 


```{r Airfares on Travel Websites }
air_df <- read.csv("airfares2_r.csv")

air_anova <- aov(Cost ~ Flight + Service, data = air_df)

anova(air_anova)

```

This analysis of variance gives us a p-value of 0.15 which is grater
than 0.05, so therefore we would fail to reject the null hypothesis the
data is not significant enough to suggest the mean prices are different 