---
title: "Chi-Square Tests with R"
author: "Ramsey EL Lethy"
date: "2023-10-13"
output:
  pdf_document: default
  html_document: default
---



```{r #Test of Equality Population Proportions of Test Independence}

#Test of Equality Population Proportions of Test Independence 

#Read our data 
auto_loyalty_df <- read.csv("autoloyalty_r.csv")

#Generate subsets for each 
impala_df <- subset(auto_loyalty_df, Automobile == "Chevrolet Impala")
fusion_df <- subset(auto_loyalty_df, Automobile == "Ford Fusion")
accord_df <- subset(auto_loyalty_df, Automobile == "Honda Accord")

#omit the invalid spaces
impala_df <- na.omit(impala_df)
fusion_df <- na.omit(fusion_df)
accord_df <- na.omit(accord_df)

#subset of the 'Yes' Responses
yes_impala <- subset(impala_df, Likely.Repurchase == "Yes")
yes_fusion <- subset(fusion_df, Likely.Repurchase == "Yes")
yes_accord <- subset(accord_df, Likely.Repurchase == "Yes")

#Nrow for the number of yes responsees
num_yes_impala <- nrow(yes_impala)
num_yes_fusion <- 
  nrow(yes_fusion)
num_yes_accord <- nrow(yes_accord)

#subset of the 'No' responses
no_impala <- subset(
  impala_df, Likely.Repurchase == "No")
no_fusion <- subset(fusion_df, Likely.Repurchase == "No")
no_accord <- subset(accord_df, Likely.Repurchase == "No")

#nrow for the number of 'no' repsonses
num_no_impala <- nrow(no_impala)
num_no_fusion <- nrow(no_fusion)
num_no_accord <- nrow(no_accord)

#sample size for each car
samp_size_impala <- num_yes_impala + num_no_impala
samp_size_fusion <- num_yes_fusion + num_no_fusion
samp_size_accord <- num_yes_accord + num_no_accord

#finally we can do our proportion test 
test <- prop.test(c(num_yes_impala, num_yes_fusion,
                    num_yes_accord), c(samp_size_impala,
                                       samp_size_fusion, samp_size_accord),
                  correct = FALSE)

list(test)



```
The Chi Squared value is 7.891 and the P-vaule 
is 0.0193, Since the p value is less than our 
alpha, we will reject the null hypothesis
there is enough evidence to suggest that 
the proportions are not equal 


```{r Chi Square Goodness-of-Fit Test}
#Chi Square goodness of Fit test

#read our file 
smr_df <- read.csv("scottmarketingsummary_r.csv")

#create a subset of each company 
company_a <- subset(smr_df, Company == "A")
company_b <- subset(smr_df, Company == "B")
company_c <- subset(smr_df, Company == "C")

#create a vector of obeserved responses 

observed = c(company_a$Count, company_b$Count, company_c$Count)

#create a vector of expected proportions
expected_props = c(0.3,0.5,0.2)

#lets run our test!
test_chi <- chisq.test(x = observed, p = expected_props)

list(test_chi)

```
We get a chi square value of 7.34 and a p value of 0.0255 
since 0.05 our alpha vaue, is greater than 0.0255 we reject 
the null hypothesis, there is enough evidence to suggest that 
the proportions are not equal to 0.3, 0.5 and 0.2 respectively. 


```{r Use of Social Media}
#Test of Equality of  Population Proportions for the Use of Social Media 

#read out data 
socialmedia_df <- read.csv("socialmedia_r.csv")

View(socialmedia_df)
#create subsets for each country
uk_df <- subset(socialmedia_df, Country == "United Kingdom")
china_df <- subset(socialmedia_df, Country == "China")
russia_df <- subset(socialmedia_df, Country == "Russia")
usa_df <- subset(socialmedia_df, Country == "United States")


#subset of the yes responses 
yes_uk <- subset(uk_df, Use_Social_Media == "Yes")
yes_china <- subset(china_df, Use_Social_Media == "Yes")
yes_russia <- subset(russia_df, Use_Social_Media == "Yes")
yes_usa <- subset(usa_df, Use_Social_Media == "Yes")

#nrow for the number of yes responses
num_yes_uk <- nrow(yes_uk)
num_yes_china <- nrow(yes_china)
num_yes_russia <- nrow(yes_russia)
num_yes_usa <- nrow(yes_usa)

#subset of the no responses
no_uk <- subset(uk_df, Use_Social_Media == "No")
no_china <- subset(china_df, Use_Social_Media == "No")
no_russia <- subset(russia_df, Use_Social_Media == "No")
no_usa <- subset(usa_df, Use_Social_Media == "No")

#nrow for the number of no responses
num_no_uk <- nrow(no_uk)
num_no_china <- nrow(no_china)
num_no_russia <- nrow(no_russia)
num_no_usa <- nrow(no_usa)

#sample size for each country
samp_size_uk <- num_no_uk + num_yes_uk
samp_size_russia <- num_no_russia + num_yes_russia
samp_size_china <- num_no_china + num_yes_china
samp_size_usa <- num_no_usa + num_yes_usa

#porportion test 

test_countries <- prop.test(c(num_yes_uk, num_yes_china,
                    num_yes_russia, num_yes_usa), c(samp_size_uk,
                                       samp_size_china, samp_size_russia,
                                       samp_size_usa),
                  correct = FALSE)

list(test_countries)


```
A. This test gives us a Chi Squared Statistic of 79.35 and a 
p-value of 0.000, since our p- value is less that our alpha values of
0.05, we reject the null hypothesis, and conclude that there isnt enough
evidence to suggest the proportions are different.

B. The sample Proportions for each country are as follows 
prop 1 prop 2 prop 3 prop 4 
0.60   0.43   0.49   0.64



```{r M&M Candy Colors}
#M&M Candy colors 

candy_df <- read.csv("mandm.csv")

blue_df <- subset(candy_df, Color == "Blue")
brown_df <- subset(candy_df, Color == "Brown")
green_df <- subset(candy_df, Color == "Green")
orange_df <- subset(candy_df, Color == "Orange")
red_df <- subset(candy_df, Color == "Red")
yellow_df <- subset(candy_df, Color == "Yellow")

observed_candy = c(105,72,89,84,70,80)

expected_props_candy = c(0.24,0.13,0.20,0.16,0.13,0.14)

test_chi_candy <- chisq.test(x = observed_candy, p = expected_props_candy)

list(test_chi_candy)

```
We get a Chi Squared Value of 5.852 and a 
P-Value of 0.3209 since our P-Value is greater
than our alpha value of 0.05, we fail to reject the null
hypothesis, there is not enough evidence to suggest
the proportions are not 0.24, 0.13, 0.20, 0.16,0.13,0.14

