---
title: "Inference About Two Populations"
author: "Ramsey EL Lethy"
date: "2023-09-23"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r exam_scores}
#Inferences about two populations with R 


exam_scores_df <- read.csv("examscores_r.csv")

#create a data frame consisting only of the variable center

center_a_df <- subset(exam_scores_df,select = Center.A)

#remove the rows with missing values 
center_a_df <- na.omit(center_a_df)

#calculate sample size

samp_size_a <- nrow(center_a_df)

#new vector of A scores
a_scores <- center_a_df$Center.A

#calculate sample mean 
samp_mean_a <- mean(a_scores)

#we do the same thing for Center B observatons
 
center_b_df <- subset(exam_scores_df,select = Center.B)

center_b_df <- na.omit(center_b_df)

samp_size_b <- nrow(center_b_df)

b_scores <- center_b_df$Center.B

samp_mean_b <- mean(b_scores)

#calculate seDifference 
std_error <- sqrt((10^2 / samp_size_a) + (10^2 / samp_size_b))

#get our test stat!

test_z <- (samp_mean_a - samp_mean_b)/std_error

#computer P-Value 
p_value <- 2*pnorm(abs(test_z),lower.tail = FALSE)

#lets see what we got!
list(test_z,p_value
     )

```
We get a Z-Stat of 1.656
Our p-value is 0.09 > 0.05 so we fail to reject the null 
## Including Plots

You can also embed plots, for example:

```{r Check Account}
check_acct_df <- read.csv("checkacct_r.csv")

#Create vectors
beechmontBalance <- check_acct_df$Beechmont

cherry_grove <- check_acct_df$Cherry.Grove

#T-TEst
test <- t.test(cherry_grove, beechmontBalance, paired = FALSE,
               alternative = "two.sided", conf.level=0.95)
#Results
list(test)

```
We get a t value of 2.954 df = 47.805
We get at p-value of 0.00048
Since this p-value is greater than .05 we reject the null 


```{r matched}
matched_df <- read.csv("matched_r.csv")
View(matched_df
     )
#Vectors!
times_1 <- matched_df$Method.1
times_2 <- matched_df$Method.2

#T Test!
test <- t.test(times_1, times_2, paired=TRUE,
              alternative = "two.sided", conf.level = 0.95)
#Results!
list(test)

```
We get a t value of 2.1958
With 5 Degree's of freedom 
We get a p- value of 0.079 > 0.05 so we fail to reject the null  


```{r tax prep}
tax_prep_df <- read.csv("taxprep_r.csv")
View(tax_prep_df
     )
office_1_df <- subset(tax_prep_df,select = Office.1)
office_1_df <- na.omit(office_1_df)
errors_1 <- subset(office_1_df,Office.1 == "Yes")
num_errors_1 <- nrow(errors_1)
not_errors_1 <- subset(office_1_df,Office.1 == "No")
num_not_errors_1 <- nrow(not_errors_1)
samp_size_1 <- num_errors_1 + num_not_errors_1


office_2_df <- subset(tax_prep_df,select = Office.2)
errors_2 <- subset(office_2_df,Office.2 == "Yes")
num_errors_2 <- nrow(errors_2)
not_errors_2 <- subset(office_2_df,Office.2 == "No")
num_not_errors_2 <- nrow(not_errors_2)
samp_size_2 <- num_errors_2 + num_not_errors_2

test <- prop.test(c(num_errors_1, num_errors_2),
                  c(samp_size_1, samp_size_2), 
                  alternative = "two.sided", conf.level = 0.95,
                  correct = FALSE)


list(test) 

```
We get a Chi-Squared Stat of 3.4084, with 1 Degree of freedom 
Our p-value is 0.064 > 0.05 so we fail to reject the null 
```{r Hotel Price Comparison}
#Hotel Price Comparison 

hotel_price_df <- read.csv("hotel_r.csv")

#define the hypothesis
#HO u1 - u2 >= 0 (mean price of hotel in atl is greater than houston)
#HA u1 - u2 < 0 (the mean price of a hotel in atl lower than mean price in hou)

#create a data fram consisting only of the variable Atlanta
atlanta_df <- subset(hotel_price_df, select = Atlanta)
atlanta_df <- na.omit(atlanta_df)

#Data frame consisting of only the variable Houston
houston_df <- subset(hotel_price_df, select = Houston)
houston_df <- na.omit(houston_df)

#Generate Sample Size
sample_size_atl <- nrow(atlanta_df)
sample_size_hou <- nrow(houston_df)

#Vector of atl and hou
atl_entries <- atlanta_df$Atlanta
hou_entries <- houston_df$Houston
 #sample mean 
sample_mean_atl <- mean(atl_entries)
sample_mean_hou <- mean(hou_entries)
#std error
std_error_ah <- sqrt((10^2 / sample_size_atl) + (10^2 / sample_size_hou))
#z-test
z_test_result <- (sample_mean_atl - sample_mean_hou) / std_error_ah
#p-value
p_value <- pnorm(z_test_result, lower.tail = TRUE)
#results!
list(test)

```
We get a Z test of 3.404
and a p value of .06 > .05 so we fail to reject the null 
```{r pressure, echo=FALSE}
#SAT Scores 

testscores_df <- read.csv("testscores_r.csv")

math_score_df <- testscores_df$Math_Score
writing_score_df <- testscores_df$Writing_Score

test <- t.test(writing_score_df, math_score_df, paired = FALSE, 
               alternative = "two.sided", conf.lvel = .99)

list(test)

```
We get a t value of -0.74, with DF = 20.13
P = 0.463 > .05 So We fail to reject the null 
Sample Estimates x = 489 y = 514
Math reports higher mean score 