---
title: "Various Hypothesis Tests"
author: "Ramsey EL Lethy"
date: "2023-09-15"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Hypothesis Testing of a Population Mean, 
Known Standard Deviation
```{r HypTest PSTD Known}

#Take in data file from working directory
golf_test_df <- read.csv("golftest_r.csv")


#Calculate the sampel size
samp_size <- nrow(golf_test_df)

#calculate the standard eorror of the sample mean 
std_error <- 12/sqrt(samp_size)

#create a vector of yards
yards <- golf_test_df$Yards

#calculate the value of the test statistic
test_z <- (mean(yards) - 295)/std_error 

#calculate the p-value
p_value <- 2*pnorm(abs(test_z),lower.tail=FALSE)

#print the resulting test stat and p-value
list(test_z,p_value)

```
Our Z Stat calculated to 1.532
Which then processed a p-value of 0.126
Since the P-value is bigger than our alpha value, we fail to reject this null hypothesis. 


Cost Of Telephone Surveys
```{r Hyptest1}
fowle_r_df <- read.csv("fowle_r.csv")

# Null Hypothesis MU <= 15
# Alt Hypothesis  MU > 15
# *Right tail Test*

#Calculate our sample size 
sample_size <- nrow(fowle_r_df)

#solve for standard error
standard_error <- 4/sqrt(sample_size)

#create a vector of time
time <- fowle_r_df$Time

#solve for the z statistic
test_z_stat <- (mean(time) - 15) / standard_error 

#solve for p value
p_value <- pnorm((test_z_stat),lower.tail=FALSE)

#print results .0015 < a = .01 (reject H0)
list(test_z_stat,p_value)

```
This test produces a Z-statistic of 2.958
and a P-value of .0015. Since .0015 < .01, we 
Will reject the null hypothesis


Hypothesis testing of a population Mean, 
UnkownStandard Deviation 
```{r Hyptest2}
#read data to r
coffee_df <- read.csv("coffee_r.csv")

#create the vector of weights
weights <- coffee_df$Weight
#run our t test
test <- t.test(weights, alternative="less", 
              mu = 3, conf.level = 0.99)
#gather results!
list(test)
```
This test with a t value of -2.823 and p-value of .0038 indicates a lower tail hypothesis test.
Since our p-value is less than our alpha value,
we will reject the null hypothesis 

Cost Of Residential Water
```{r Hyptest3}

#set our vairable to the data set
water_df <- read.csv("water_r.csv")

#we're testing to see if other U.S Public
#Utilities differ than Tulsa's Rate, 
#So we will use a two tail test

# mu1 = mu2 (mu1 - mu2 = 0)
# mu1 != mu2 (mu1 - mu2 != 0)
str(water_df)
#gather our vector
water <- water_df$Rate
#run our test!
test <- t.test(water, alternative="two.sided", mu=21.62
               ,conf.level = 0.95)
#lets see our results
list(test)
```
Our T value came out to -1.1481, with 41 degrees of freedom, and a p-value of 0.2576
comparing that to our alpha value of 0.05, we see that 0.2576 > 0.05, so we fail to reject the null. There is not enough evidence to suggest the true mean is not equal to 21.62
Our sample estimated the mean of x to be 20.2383, using the critical value approach, if our value to test (21.62) >= to our sample mean, we will fail to reject the Null Hypothesis. (Which it is)

Hypothesis Testing of a Population Proportion
```{r Hyptest4}
women_golf_df <- read.csv("womengolf_r.csv")

#check our sample size
samp_size <- nrow(women_golf_df)
#creating a data frame consisting only of the Female responses
females <- subset(women_golf_df,Golfer=="Female")
#count the number of female responses
female_count <- nrow(females)
#run our test!
test <- prop.test(female_count,samp_size,p=0.20,
                  alternative="greater",conf.level = 0.95,
                  correct=FALSE)
list(test)

```
This test calculates X-squared to be 6.25
df = 1
p-value = 0.00621
since 0.00621 < 0.05 we will reject the null 

Coupon Usage
```{r Hyptest5}
eagle_df <- read.csv("eagle_r.csv")
#we will check the number of "yes" responses
count_yes <- sum(eagle_df$Used_Coupon == "Yes")
#check total our observation
total_obs <- nrow(eagle_df)
#gather a sample of yes answers
sample_proportion <- count_yes / total_obs
#run our test!
result <- prop.test(count_yes, total_obs, p = 0.1, alternative = "greater")

list(result)
```
This test calculates X-squared to be 0.6944
df = 1
p-value = 0.2023
since 0.2023 > 0.05 we fail to reject the null
We do not have enough evidence to conclude the
proportion of users is greater than 10%