---
title: "Various Linear Regression Models"
author: "Ramsey EL Lethy"
date: "2023-11-04"
output:
  pdf_document: default
  html_document: default
---
```{r Passing and Winning in the NFL }
#Passign and Winning in the NFL 
library(ggplot2)
passing_df <- read.csv("nflpassing_r.csv")

str(passing_df
    )

# Create the plot
my_plot <- ggplot(passing_df, aes(x = Yds, y = Win)) +
  geom_point() +
  labs(x = "Yds", y = "Win", title = "Scatterplot of Yds vs. Win")

# Display the plot in a separate window
plot(my_plot)

passing_SLR <- lm(Win ~ Yds, data = passing_df)
new_data <- data.frame(Yds = 6.2)

summary(passing_SLR)

prediction <- predict(passing_SLR, newdata = new_data)

```
since our ploints are upward sloping
this scatter diagram shows a positive relationship
between our independant variable, Yards and our 
dependant variable Wins

the slope of our regression line is positive 
representing a positive relationship between x and y 
our regression equation is -70.391 + 17.175x 

our prediction expects the cheifs to have about 36 wins 
```{r Age and the Price of Wine}
wine_df <- read.csv("wineprices_r.csv")

wine_plot <- ggplot(wine_df, aes(x = Age, y = Price)) +
  geom_point() +
  labs(x = "Age", y = "Price", title = "Scatterplot of Age vs. Price")

# Display the plot in a separate window
plot(wine_plot)

wine_LR <- lm(Price ~ Age, data = wine_df)

summary(wine_LR)


```
this scatter diagram 
shows a positive relationship
between x (age) and y (price)


This regressin shows a slope of 6.9520, 
this slope is positive
suggesting a upward sloping or positive relationship between 
the age of the wine and its price
```{r Price and Weight of Bicycles}

bike_df <- read.csv("racingbicycles_r.csv")

bike_r <- lm(Price ~ Weight, data = bike_df)

summary(bike_r)

newdata_bike = data.frame(Weight = 15)
predict(bike_r,newdata = newdata_bike)
```
R Squared came to be 0.86
Since R > 0.5, we can say that 
the estimated regression equation 
provides a good fit of the data 
86 percent of the data can be attributed to our regression 

at a weight of 15 pounds 
our regression predicts the price to be $6980.006

```{r College GPA and Salary }
college_df <- read.csv("gpasalary_r.csv")

college_plot <- ggplot(college_df, aes(x = GPA, y = Salary)) +
  geom_point() +
  labs(x = "GPA", y = "Salary", title = "Scatterplot of GPA vs. Salary")

# Display the plot in a separate window
plot(college_plot)

college_lr <- lm(Salary ~ GPA, data = college_df)

summary(college_lr)
```
this scatter diagram 
shows a positive relationship
between x (GPA) and y (Salary)

Regression Equation Salary = -110210 + 72990 * GPA

At a significance level of 0.05, our f statistic 
is 104.3, giving a p-value to be 0.000
since p < 0.05 this shows a significant relationship between
GPA and salary 10 years after college
we reject the null hypothesis and say that 
GPA has some effect on salary 10 years after college 



```{r Entertainment Spending}
#Entertainment Spend 

spending_df <- read.csv("businesstravel_r.csv")

spending_lr <- lm(Entertainment ~ Room.Rate, data = spending_df)


spending_function <- function(x) {
  y <- 17.49 + 1.0334 * x 
  return(y)
}


prediction <- spending_function(89)

print(prediction)

newdata_spending = data.frame(Room.Rate = 89)
predict(spending_lr,newdata_spending,interval="confidence", level = 0.95)

newdata_prediction = data.frame(Room.Rate = 128)
predict(spending_lr, newdata_prediction, interva = "prediction", level = 0.95 )

```
we predict the value of y to be 109.4626

confidence interval stats lower 94.84742  upper 124.0833

prediciton interval stats lower 110.6927 upper 188.8444
